Cover letter for the original IJRR submission

### 2017-04-19 body/abstract.tex
corrections fro mrev 1

% !TEX root =  ../main.tex
We present a contact planning framework for complex legged locomotion tasks: standing up, climbing stairs using a handrail, crossing rubble and getting out of a car. The need for such a planner was highlighted at the Darpa Robotics Challenge, where multi-contact behaviors
could not be demonstrated.

>>>>

% !TEX root =  ../main.tex
We present a contact planner for complex legged locomotion tasks: standing up, climbing stairs using a handrail, crossing rubble and getting out of a car. The need for such a planner was shown at the Darpa Robotics Challenge, where such behaviors
could not be demonstrated (except for egress).

### 2017-04-19 body/conclusion.tex
corrections fro mrev 1

\subsection{Addressing $\mathcal{P}_3$}
This paper focuses on contact planning of multi-contact locomotion. While it does not address directly the continuous motion generation,
we had to prove that the plans produced by our planner can be executed by the considered robots. Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory},
and most existing solutions do not consider collisions with the environment during the planning, making it hard to validate constrained problems such as the polaris scenario
or the table egress in \cite{DBLP:conf/iser/EscandeKMG08}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot. Despite the fact that our method does not consider joint torque limits in the planning
phase, the robot is able to execute our plans, achieving to demonstrate the interest of our framework.
%~ \adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

>>>>

\subsection{Addressing $\mathcal{P}_3$}
This paper focuses on contact planning of multi-contact locomotion. While it does not address directly the continuous motion generation,
we had to prove that the plans produced by our planner can be executed by the considered robots. Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory}, for which the implemented a solution in this paper, described in appendix 
%~ and most existing solutions do not consider collisions with the environment during the planning, making it hard to validate constrained problems such as the polaris scenario
%~ or the table egress in \cite{DBLP:conf/iser/EscandeKMG08}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot.
In these scenarios the robot is able to execute our plans, while respecting the constraints inherent to the problem: joint kinematics and torque limits and
collision avoidance.
%~ \adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

### 2017-04-19 body/overview.tex
corrections fro mrev 1

\medskip
\textbf{A robot} is a kinematic chain $R$ composed of: a root $R^0$, and $l$ limbs $R^k, 1 \leq k \leq l$, attached to the root.
The root has $r \geq 6$ DoFs: for instance, HRP-2 has two extra DoFs in the torso, such that we have $r=8$.
Therefore $R$ is fully described by a configuration $\mathbf{q} \in  \mathbb{R}^{r+n}$, with $n$ the number of joint DoFs.
$\mathbf{q}$ is decomposed as follows:
\begin{itemize}
	\item $\mathbf{q}^k$ is a configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ is a vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ is the world coordinates vector of $R^0$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
\noindent $W^0$ is a volume encompassing $R^0$ (Figure~\ref{fig:HyQ_roms}). \\
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 $W^k$ is the reachable workspace of a limb $R^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{j.lim}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position (in the root frame) of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{j.lim}$ is the space
of admissible limb joint configurations. We also define \mbox{$W = \bigcup_{k=1}^{l}W^k$}, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ for the root configuration $\mathbf{q}^{0}$.

>>>>

\medskip
\textbf{A robot} is a kinematic tree $R$ composed of: a root $R^0$, and $l$ limbs $R^k, 1 \leq k \leq l$, attached to the root.
The root has $r \geq 6$ DoFs: for instance, HRP-2 has two extra DoFs in the torso, such that we have $r=8$.
Therefore $R$ is fully described by a configuration $\mathbf{q} \in  \mathbb{R}^{r+n}$, with $n$ the number of joint DoFs.
$\mathbf{q}$ is decomposed as follows:
\begin{itemize}
	\item $\mathbf{q}^k$ is a configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ is a vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ is the world coordinates vector of $R^0$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
\noindent $W^0$ is a volume encompassing $R^0$ (Figure~\ref{fig:HyQ_roms}), or equal to it. $W^0$ is typically a low-polygonal bounding shape
of $R^0$ for performance reasons.\\
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 $W^k$ is the reachable workspace of a limb $R^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{j.lim}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position (in the root frame) of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{j.lim}$ is the space
of admissible limb joint configurations. We also define \mbox{$W = \bigcup_{k=1}^{l}W^k$}.

Each $W^i, 0 \leq i \leq l$ is  attached to one joint of the root, such that 
$W^i(\mathbf{q}^{0})$ is the volume $W^i$ transformed the by root configuration $\mathbf{q}^{0}$.

### 2017-04-19 body/rbprm.tex
cosmetics
corrections fro mrev 1

% !TEX root =  ../main.tex
%~ \section{Computing a guide path in $C^0_{reach}$ ($\mathcal{P}_1$) }
\section{Root path planning in the \textit{contact reachable space}}
\label{rbprm}

@@

\textbf{necessary conditions:}
For a contact to be possible, an obstacle $O_i \subset O$ necessarily intersects the reachable workspace $W(\mathbf{q}^{0})$ of the robot (Figure~\ref{fig:contact_gen}--1). Also the torso of the robot $W^0(\mathbf{q}^{0})$ must necessarily be collision-free.
Therefore we can define an outer approximation  $ \mathcal{C}^0_{\textrm{\it Nec}} \supset$ $C_{Contact}^0$ as: 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Nec}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%~ The inclusion $C_{Contact}^0$ $\subset \mathcal{C}^0_{\textrm{\it Nec}}$ is straightforward, but it is very important.
%~ Planning in $\mathcal{C}^0_{\textrm{\it Nec}}$ allows a strong reduction of the search space, while not discarding 
%~ any root configuration that could lead to a solution.

@@

\textbf{sufficient condition:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

@@

%~ However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. 
We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
 \label{eq:reachc}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
The parametrization of $s$ defines a trade-off:
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$.
 %~ We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient, but less and less necessary.  
Eq.~\ref{eq:reachc} thus defines the \textit{reachability condition}. We fix a value $s^*$ for $s$ and define  $C_{Reach}^0 = C^0_{s^*}$.
The computation of $s^*$ is detailed in Section~\ref{sec:params}. 
In Appendix~\ref{app:rom}, we give a generic method to compute the $W$ volumes appearing in the definition of $C_{Reach}^0$.

\subsection{Computing the guide path in $C_{reach}^0$}
$C_{Reach}^0$ can be sampled efficiently thanks to Eq.~\ref{eq:reachc}, and can thus be used with a standard motion planner.
The only significant change is to replace the collision checking with the \textit{reachability condition}.
Our current implementation of these modifications is based on the Bi-RRT planner \citep{770022} provided by the HPP software~\citep{7759083}.

@@

\subsubsection*{}
This Section has presented a guide path planner for the geometric root of a robot. This planner is implemented as a low-dimensional sampling-based 
algorithm. Given start and goal configurations, it outputs a continuous path for the root of the robot. 
%~ \deladp{Thanks to the \textit{reachability condition}, we assume that for any configuration in the path there exists a joint configuration that results in static equilibrium.}
%~ Thanks to these modifications, the problem of planning a \gls{contact reachable} path is reduced to a geometric collision-checking problem, of low dimension (6 for HyQ, 8 for HRP-2 that has 2 joints in the torso).
%~ By doing this we can solve the problem with a sampling-based approach, in \gls{interactive} computation times.

>>>>

% !TEX root =  ../main.tex
%~ \section{Computing a guide path in $C^0_{reach}$ ($\mathcal{P}_1$) }
\section{Root path planning in the contact reachable space}
\label{rbprm}

@@

\textbf{necessary conditions:}
For a contact to be possible, an obstacle $O_i \subset O$ necessarily intersects the reachable workspace $W(\mathbf{q}^{0})$ of the robot (Figure~\ref{fig:contact_gen}--1). Also the torso of the robot $W^0(\mathbf{q}^{0})$ must necessarily be collision-free.
Therefore we can define an outer approximation  $ \mathcal{C}^0_{\textrm{\it Nec}} \supset$ $C_{Contact}^0$ as: 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Nec}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%~ The inclusion $C_{Contact}^0$ $\subset \mathcal{C}^0_{\textrm{\it Nec}}$ is straightforward, but it is very important.
%~ Planning in $\mathcal{C}^0_{\textrm{\it Nec}}$ allows a strong reduction of the search space, while not discarding 
%~ any root configuration that could lead to a solution.

@@

\textbf{sufficient condition:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

@@

%~ However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. 
We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
 \label{eq:reachc}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \textrm{ \textbf{and}}\ W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
The parametrization of $s$ defines a trade-off:
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$.
 %~ We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient, but less and less necessary.  
Eq.~\ref{eq:reachc} thus defines the \textit{reachability condition}. We fix a value $s^*$ for $s$ and define  $C_{Reach}^0 = C^0_{s^*}$.
The computation of $s^*$ is detailed in Section~\ref{sec:params}. 
In Appendix~\ref{app:rom}, we give a generic method to compute the $W$ volumes appearing in the definition of $C_{Reach}^0$.

\subsection{Computing the guide path in $C_{reach}^0$}
$C_{Reach}^0$ can be sampled efficiently thanks to Eq.~\ref{eq:reachc}, and can thus be used with a standard motion planner.
The only significant change is to replace the collision checking with the \textit{reachability condition}.
Our current implementation of these modifications is based on the Bi-RRT planner \citep{770022} provided by the HPP software~\citep{7759083}. \\

@@

This Section has presented a guide path planner for the geometric root of a robot. This planner is implemented as a low-dimensional sampling-based 
algorithm. Given start and goal configurations, it outputs a continuous path for the root of the robot. 
%~ \deladp{Thanks to the \textit{reachability condition}, we assume that for any configuration in the path there exists a joint configuration that results in static equilibrium.}
%~ Thanks to these modifications, the problem of planning a \gls{contact reachable} path is reduced to a geometric collision-checking problem, of low dimension (6 for HyQ, 8 for HRP-2 that has 2 joints in the torso).
%~ By doing this we can solve the problem with a sampling-based approach, in \gls{interactive} computation times.
